# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
activityData <- read.csv("activity.csv")
head(activityData)
```

## What is mean and median of total number of steps taken per day?
### First Plotting the total number of steps per day against dates
```{r}
totalStepsByDate<-tapply(activityData$steps, activityData$date, FUN=sum, na.rm=TRUE)
barplot(totalStepsByDate, xlab="Date", ylab="Total Number of Steps")
png(filename="TotalStepsVsDate")
```
### Now finding the mean and the median of the total number of steps taken per day
```{r}
averageNumberOfStepsPerDay <- mean(totalStepsByDate)
averageNumberOfStepsPerDay
medianNumberOfStepsPerDay <- median(totalStepsByDate)
medianNumberOfStepsPerDay
```

## What is the average daily activity pattern?
### First plotting the average number of steps by period
```{r}
averageNumberOfStepsPerPeriod <- tapply(activityData$steps, INDEX=activityData$interval, FUN=mean, na.rm=TRUE)

plot(names(averageNumberOfStepsPerPeriod), averageNumberOfStepsPerPeriod, xlab="Interval", ylab="Average Number of Steps")

lines(names(averageNumberOfStepsPerPeriod), averageNumberOfStepsPerPeriod)

# Now plotting a time series Histogram for a different visualization
barplot(averageNumberOfStepsPerPeriod)
```
### Now calculating the period with the greatest number of steps
```{r}
periodWithHighestAvgSteps <- names(which.max(averageNumberOfStepsPerPeriod))
periodWithHighestAvgSteps
```

## Imputing missing values

### Calculating the number of records with missing "steps"" values
```{r}
moOfRecordsWithMissingSteps <- nrow(activityData) - nrow(activityData[complete.cases(activityData),])
moOfRecordsWithMissingSteps
```
### Now imputing the missing values using the strategy defined as follows:
### The missing "steps" value for a specific period is the average (mean) value of all the "steps" for that particular period
```{r}
activityDataWithNARemoved <- data.frame()
activityDataWithNAFilled <- data.frame()
library(plyr)
activityDataWithNaRemoved <- activityData[which(activityData$steps != "NA"), ]
avg_by_interval <- ddply(activityDataWithNaRemoved, .(interval), summarise, steps=mean(steps))
activityDataWithNAFilled <- merge(x=activityData, y=avg_by_interval, by.x="interval", by.y="interval", all=TRUE)
activityDataWithNAFilled <- arrange(activityDataWithNAFilled, date)
for(i in 1:nrow(x=activityDataWithNAFilled)){
    record <- activityDataWithNAFilled[i,]
    if(is.na(x=record$steps.x)){
        activityDataWithNAFilled[i,]$steps.x <- activityDataWithNAFilled[i,]$steps.y
    }
}

```
### Now creating a histo gram using the data with no NAs
```{r}
modifiedTotalStepsByDate<-tapply(activityDataWithNAFilled$steps.x, activityDataWithNAFilled$date, FUN=sum)
barplot(modifiedTotalStepsByDate, xlab="Date", ylab="Total Number of Steps")
```
### Now calculating the mean and median with the new set of data
```{r}
averageNumberOfStepsPerDay <- mean(modifiedTotalStepsByDate)
averageNumberOfStepsPerDay
medianNumberOfStepsPerDay <- median(modifiedTotalStepsByDate)
medianNumberOfStepsPerDay
```

## Are there differences in activity patterns between weekdays and weekends?
